---
title: "PA1_StatisticalInference"
author: "Rafael Mendez"
date: "August 19, 2015"
output: pdf_document
---
Purpose: 
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.  

Load in Global Knitr Parameters
```{r}
library(knitr)
opts_chunk$set(echo = TRUE, results = 'hold')  
```

#1. Show the sample mean and compare it to the theoretical (expected) mean of the distribution.

##Create Simulation with 40 samples and 1000 num of simulations 
```{r}
#Set the constants
set.seed(50) # Make simulation reproducible
lambda <- 0.2  # set for rex, the rate parameter
samples <- 40 #number of exponetials being distributed
sim <- 1000 # number of tests
sample_mean <- NULL #initalize sample_mean
for (i in 1:sim) sample_mean <- c(sample_mean, mean(rexp(samples, lambda)))
```

##Compare Sample Mean vs Theoritical Mean 

##Report the Sample Mean
```{r}
mean_of_means<- round(mean(sample_mean))
mean_of_means
```

##Report the Theoretical Mean - expected mean mu of expoential distribution rate lambda given = .20
```{r}
mu <- 1/lambda 
mu
```

results: Shows that the sample mean and expected mean are closely related.

#2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

##Compare Sample Variance vs Theoretical Variance

##Report the Sample Variance
```{r}
#var() function to get sample variance 
variance_sample<-var(sample_mean)
variance_sample
```

##Calculate the Expected Standard Deviation of a exponential distribution of rate parameter lambda 
As variance = sd ^ 2. Find the sd through 1/(lambda/sqrt(#ofsamples))
```{r}
sd <- 1/lambda/sqrt(samples)
sd
```

##Report the Theoretical Variance of Distribution
```{r}
variance_theoretical<- sd^2
variance_theoretical
```

#3. Show that the distribution is approximately normal.
We use a normal distribution of expected values to compare the population means and standard deviation.
The normal fitted curve shows the normal distribution
 
```{r echo=FALSE}
#Note that histograms are strongly affected by the number of bins used
normal_curve_hist <- hist(sample_mean, breaks=30, col = "red") 
#Add a normal curve
xfit <- seq(min(sample_mean), max(sample_mean), length = 40) #dictates 
yfit <- dnorm(xfit, mean = mean(sample_mean), sd = sd(sample_mean)) 
yfit <- yfit  * diff(normal_curve_hist$mids[1:2]) * length(sample_mean) 
lines(xfit, yfit, lwd=3)
```

